Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-15 introduction.tex
Move to local

A complexity theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks the possibility for an efficient classical verifier (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations. 
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,PR:FitKas17,mf16}) considers the case where the $\BPP$ verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non communicating quantum provers. 
In a recent breakthrough, Mahadev~\cite{mahadev_2018} proposed the first protocol of classical verification of quantum computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines. 
The technique invented therein has inspired many  subsequent developments of CVQC protocols with improved parameters and functionality (e.g.~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}). 
We refer curious readers to the survey~\cite{survey:GKK19} for details.

With the newly developed techniques, we revisit the classical verification of quantum computation problem from both a philosophical and a practical point of view. 
Our first observation is that the sampling version of $\BQP$ (e.g., as formulated by Aaronson~\cite{aaronson_2013})

>>>>

A complexity theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks the possibility for an efficient classical verifier (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations. 
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,PR:FitKas17,mf16}) considers the case where the $\BPP$ verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non communicating quantum provers. 
In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of classical verification of quantum computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines. 
The technique invented therein has inspired many  subsequent developments of CVQC protocols with improved parameters and functionality (e.g.~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}). 
We refer curious readers to the survey~\cite{survey:GKK19} for details.

With the newly developed techniques, we revisit the classical verification of quantum computation problem from both a philosophical and a practical point of view. 
Our first observation is that the sampling version of $\BQP$ (e.g., the $\SampBQP$ formulated by Aaronson~\cite{aaronson_2013}) might be a more appropriate notion to serve the purpose of the original problem. 
Philosophically, experiments of quantum mechanics usually generate samples and statistical information, which is well demonstrated in the famous double-slit experiment. 
Moreover, a lot of quantum algorithms from the famous Shor's~\cite{Shor} and Grover's~\cite{Grover} algorithms to some recent developments in machine learning and optimization (e.g.~\cite{brando_et_al:LIPIcs:2019:10603, AGGW17,pmlr-v97-li19b}) contain a significant quantum sampling component. 
The fact that almost all quantum supremacy tasks (e.g.,~\cite{Boson, IQP, nature-google}) are sampling ones strengthens the importance of $\SampBQP$. 
Even though the relation between $\BQP$ and $\SampBQP$ is relatively understood in the plain model, what are the potential technical challenges raised by their subtle differences in the context of CVQC and how to resolve them is however far from clear. 
In particular, it is a prior unclear whether it is feasible to develop a CVQC protocol for $\SampBQP$ based on Mahadev's technique~\cite{FOCS:Mahadev18a}.

### 2020-04-15 qpip0.tex
Move to local

In this section, we combine \cref{QPIP1thm} with results from \cite{FOCS:Mahadev18a} to create a delegation protocol for $\SampBQP$ for fully classical clients. To boost the soundness, we need  polynomial \Ethan{Wording is strange/misleading} rounds, so we use results from \cite{parallelrep} to run polynomially many copies in parallel and get a constant-round protocol.

@@

We prove by contradiction. Suppose that
$$D_3\not\approx_{c, \eps+\eps'} D_5,$$
then there exist a quantum circuit $A$ of size $poly(n)$ such that
$$\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| > \eps+\eps'.$$
Since $$\norm{D_4 - D_6}_q \leq \eps'$$, we know that
$$\L\|\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \leq \eps'.$$
Since $D_1=D_3+D_4$ and $D_2=D_5+D_6$, we have that
$$ \Pr[A({D_1})=1]= \Pr[A({D_3})=1]+ \Pr[A({D_4})=1]$$
$$ \Pr[A({D_2})=1]= \Pr[A({D_5})=1]+ \Pr[A({D_6})=1].$$
So
\begin{align}
    &\L\|\Pr[A({D_1})=1]-\Pr[A({D_2})=1]  \R\| \nn \\
    =&   \L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1] +\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \nn \\
    \geq  &\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| - \L\|\Pr[A({D_6})=1]-\Pr[A({D_4})=1]  \R\|  \nn \\
     > &\eps
\end{align}
	This contradicts with the assumption that $D_1\approx_{c,\eps}D_2$.
\end{proof}

@@

\begin{proof}
For every cheating prover $\PNaiveStar$ which aborts, consider an alternative prover $P_{alt}$ which is $\PNaiveStar$ with the following modification: on the second round of Protocol~\ref{proto:urmila4}, if $\PNaiveStar$ would abort, $P_{alt}$ switch to run the trivial strategy specified by Fact~\ref{fact:perfectly_pass_test} instead. This is possible because the message send by the verifier on the first round is classical, so the prover can recalculate everything.

By Fact~\ref{fact:perfectly_pass_test} and the assumption of the lemma, $p_{alt}$ passes the testing round with $(1-\negl(\lambda))$ probability. Therefore by Lemma~\ref{lem:urmila-binding}, there exist a state $\rho_{alt}$ such that 
\begin{align}
    z_{commit,alt} \approx~{c,\eps} M_{XZ}(\rho_{alt},h)
\end{align}

@@

  Note that $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ has failure probabilities, and this is reflected by the fact that $\ket{\psi_{0,i,\gamma}}_{\regX,\regZ}$ and $\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}$ are  sub-normalized. $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ depend on $(U_0,U)$ and $pk,y$. \hannote{check}
  \Ethan{check whether this should be some single pk and y or the entire tuple}

@@

We also need the following technical lemma.
\begin{lemma}\label{lem:samp-tech}
For any state $\ket{\psi}$,  $\ket{\phi}$ and projectors $\{P_z\}$ such that $\sum_z P_z \leq Id$ and $P_z P_{z'} =0 $ for all $z\neq z'$, we have
$$  \sum_z |\vev{\psi|P_z|\phi}| \leq \norm{\psi}\norm{\phi} $$
\end{lemma}
\begin{proof}
\begin{align}
    \sum_z |\vev{\psi|P_z|\phi}| =&\sum_z|\vev{\psi|P_zP_z|\phi}| \nn \\
    \leq& \sum_z \norm{\bra{\psi}P_z} \norm{ P_z\ket{\phi}}  \nn \\ \leq&  \sqrt{\sum_z \norm{P_z\ket{\psi}}^2} \sqrt{\norm{ P_z\ket{\phi}}^2}     \nn \\
    \leq& \sqrt{\norm{\sum_z P_z\ket{\psi}}^2 } \sqrt{\norm{\sum_z P_z\ket{\phi}}^2 } \nn \\
    \leq & \norm{\ket{\psi}}\norm{\ket{\phi}},
\end{align}
where we used Cauchy's inequality on the second and third line and $P_z P_{z'} =0 $ on the fourth line.
\Ethan{TODO There are typos in the proof}
\end{proof}

@@

$$z_i\defeq \E_{pk,y} \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$
where $\ket{e_i,\psi}=\ket{e_i}_\regC\ket{\psi}_{\regX,\regZ}$ and $U$ is the unitary the prover applies on the last round. Note that we have averaged over $pk, y$ where as previously everything has fixed $pk$ and $y$. \hannote{add description about  averaging $pk$ here.}

Since $\{c\}$ is drawn  from $e_i$ with uniformly random $i\in [m]$, we have 
$$ z=\frac{1}{m} \sum_i z_i= \frac{1}{m} \sum_i \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$ 
	where we represent the random variable $z$ as a real non-negative diagonal matrix, with the matrix entries begin probabilities. Note that $z$ is "sub-normalized", i.e. $\tr (z)= \text{probability of getting accepted} \leq 1$.

@@

The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then picks a uniformly random $\gamma$ and  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, there are four possibilities depending on $c$ and whether the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$. If $c=0$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. If $c=1$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U$ and check whether every copy except the $i$-th copy would be accepted. If all $m-1$ copies acceptes, he reply with the $i$-th copy, otherwise reply $\bot$. If $c-0$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U_{trivial}$ and repies. If $c=1$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he replies with $\bot$. \footnote{Recall that $c=0$ is the testing round and $c=1$ is the Hadamard round.}

@@

\begin{align}
    (d,z_{good})\approx_{c, O(\eps)}(d,z_{ideal}).
\end{align}
\hannote{do we lose a factor of $m$?}
Recall that by Eq~\ref{eq:z-z-good} we have $\norm{z-z_{good}}_1 \leq \eps$ if we pick $m,T,\gamma_0=???$ \hannote{tune eps}, which implies 
\begin{align}
    (d,z)\approx_{c, O(\eps)}(d,z_{good}).
\end{align}

Therefore by Lemma~\ref{lem:computational-triangle0}, we have  
\begin{align}
     (d,z)\approx_{c, O(\eps)}(d,z_{ideal}).
\end{align}
\hannote{the composition of z ideal and triangle ineq is sketchy lol. Also need to specify different z ideal}

>>>>

In this section, we combine \Cref{QPIP1thm} with results from \cite{FOCS:Mahadev18a} to create a delegation protocol for $\SampBQP$ for fully classical clients. To boost its soundness, we need  polynomially many repetitions, so we use results from \cite{parallelrep} to run polynomially many copies in parallel and get a constant-round protocol.

@@

We prove by contradiction. Suppose that
$$D_3\not\approx_{c, \eps+\eps'} D_5,$$
then there exist a quantum circuit $A$ of size $poly(n)$ such that
$$\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| > \eps+\eps'.$$
Since $$D_4\approx_{c, \eps'}D_6,$$ we know that
$$\L\|\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \leq \eps'.$$
Since $D_1=D_3+D_4$ and $D_2=D_5+D_6$, we have that
$$ \Pr[A({D_1})=1]= \Pr[A({D_3})=1]+ \Pr[A({D_4})=1]$$
$$ \Pr[A({D_2})=1]= \Pr[A({D_5})=1]+ \Pr[A({D_6})=1].$$
So
\begin{align}
    &\L\|\Pr[A({D_1})=1]-\Pr[A({D_2})=1]  \R\| \nn \\
    =&   \L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1] +\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \nn \\
    \geq  &\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| - \L\|\Pr[A({D_6})=1]-\Pr[A({D_4})=1]  \R\|  \nn \\
     > &\eps
\end{align}
	This contradicts with the assumption that $D_1\approx_{c,\eps}D_2$.
\end{proof}

@@

\begin{proof}
For every cheating prover $\PNaiveStar$ which aborts, consider an alternative prover $P_{alt}$ which is $\PNaiveStar$ with the following modification: on the second round of $\PiMeasure$, if $\PNaiveStar$ would abort, $P_{alt}$ switch to run the trivial strategy specified by Fact~\ref{fact:perfectly_pass_test} instead. This is possible because the message send by the verifier on the first round is classical, so the prover can recalculate everything.

By \Cref{lem:trivial-4-round-strategy} and the assumption of the lemma, $p_{alt}$ passes the testing round with $(1-\negl(\lambda))$ probability. Therefore by Lemma~\ref{lem:urmila-binding}, there exist a state $\rho_{alt}$ such that 
\begin{align} \label{eq:z-alt-rho}
    z_{commit,alt} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{alt},h).
\end{align}

Note that $z_{commit,alt} = z_{commit}+z_{trivial}$, where $z_{commit}$ and $z_{trivial}$ are sub-normalized states correspond to whether $p_{alt}$ follows $\PNaiveStar$ or the trivial strategy.

By Fact~\ref{lem:trivial-4-round-strategy} and Lemma~\ref{lem:urmila-binding}, there exist a sub-normalized state $\rho_{trivial}$ such that

\begin{align} \label{eq:z-triv-rho}
    z_{trivial} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{trivial},h).
\end{align}

By Equation~\ref{eq:z-alt-rho}, Equation~\ref{eq:z-triv-rho}, and Lemma~\ref{lem:computational-triangle}, we have

\begin{align} \label{eq:z-triv-rho}
    z_{commit} \approx_{c,\negl(\lambda)} M_{XZ}(\rho_{commit},h),
\end{align}

where $\rho_{commit} \defeq \rho_{alt}-\rho_{trivial}$.\hannote{wrong, not a state.} Note that

@@

  Note that $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ has failure probabilities, and this is reflected by the fact that $\ket{\psi_{0,i,\gamma}}_{\regX,\regZ}$ and $\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}$ are  sub-normalized. $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ depend on $(U_0,U)$ and $\vec{pk},\vec{y}$. \hannote{check}

@@

We also need the following technical lemma.
\begin{lemma}\label{lem:samp-tech}
For any state $\ket{\psi}$,  $\ket{\phi}$ and projectors $\{P_z\}$ such that $\sum_z P_z \leq Id$ and $P_z P_{z'} =0 $ for all $z\neq z'$, we have
$$  \sum_z |\vev{\psi|P_z|\phi}| \leq \norm{\psi}\norm{\phi} $$
\end{lemma}
\begin{proof}
\begin{align}
    \sum_z |\vev{\psi|P_z|\phi}| =&\sum_z|\vev{\psi|P_zP_z|\phi}| \nn \\
    \leq& \sum_z \norm{\bra{\psi}P_z} \norm{ P_z\ket{\phi}} \nn \\
    \leq&  \sqrt{\sum_z \norm{P_z\ket{\psi}}^2} \sqrt{\sum_z\norm{P_z\ket{\phi}}^2} \nn \\
    \leq& \sqrt{\norm{\sum_z P_z\ket{\psi}}^2 } \sqrt{\norm{\sum_z P_z\ket{\phi}}^2 } \nn \\
    \leq & \norm{\ket{\psi}}\norm{\ket{\phi}},
\end{align}
where we used Cauchy's inequality on the second and third line and $P_z P_{z'} =0 $ on the fourth line.
\end{proof}

@@

$$z_i\defeq \E_{pk,y} \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$
where $\ket{e_i,\psi}=\ket{e_i}_\regC\ket{\psi}_{\regX,\regZ}$ and $U$ is the unitary the prover applies on the last round. Note that we have averaged over $\vec{pk}, \vec{y}$ where as previously everything has fixed $\vec{pk}$ and $\vec{y}$.

Since $\vec{c}$ is drawn  from $e_i$ with uniformly random $i\in [m]$, we have 
$$ z=\frac{1}{m} \sum_i z_i= \frac{1}{m} \sum_i \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$ 
	where we represent the random variable $z$ as a real non-negative diagonal matrix, with the matrix entries begin probabilities. Note that $z$ is ``sub-normalized", i.e. $\tr (z)= \text{probability of getting accepted} \leq 1$.

@@

The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then picks a uniformly random $\gamma$ and  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, there are four possibilities depending on $c$ and whether the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$. If $c=0$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. If $c=1$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U$ and check whether every copy except the $i$-th copy would be accepted. If all $m-1$ copies accepts, he reply with the $i$-th copy, otherwise reply $\bot$. If $c=0$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U_{trivial}$ and repies. If $c=1$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he replies with $\bot$. \footnote{Recall that $c=0$ is the testing round and $c=1$ is the Hadamard round.}

@@

\begin{align}
    (d,z_{good})\approx_{c, O(\eps)}(d,z_{ideal}).
\end{align}
% \hannote{do we lose a factor of $m$?}
Recall that by Eq~\ref{eq:z-z-good} we have $\norm{z-z_{good}}_1 \leq \eps$ if we pick $m=O(1/\eps^2), T=O(1/\eps^2),\gamma_0=\eps^8$, which implies 
\begin{align}
    (d,z)\approx_{c, O(\eps)}(d,z_{good}).
\end{align}

Therefore  we have  
\begin{align}
     (d,z)\approx_{c, O(\eps)}(d,z_{ideal}).
\end{align}
%\hannote{the composition of z ideal and triangle ineq is sketchy lol. Also need to specify different z ideal}
